---
title: "Homework 8 - SDS315"
author: "Adam Jose - avj423"
date: "2025-04-014 - \n\n https://github.com/adamvj/SDS315"
output: pdf_document
---

  
  
```{r setup, include=FALSE}
# Global options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, tidy = TRUE, fig.width=8, fig.height=6)
```

# Problem 1

### Part A
```{r, results= FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(mosaic)
creatinine = read.csv("creatinine.csv")

creat_model = lm(creatclear ~ age, data = creatinine)
summary(creat_model)

pcreat = 147.8129 -0.6198*(55)
print(pcreat)
```
We should expect a creatinine clearance rate of 113.7239 mL/min. After doing the linear regression model, we get the y intercept at 147.8129 and the slope as -0.6198. So the formula, if the age is 55, is: 147.8129 -0.6198*(55).

### Part B

After creating the linear regression model in part A, we can see in the summary that the slope is -0.6198 mL/min.

### Part C

```{r, fig.height=4, fig.width=5}
ggplot(creatinine) +
  geom_point(aes(x=age, y=creatclear)) +
  geom_smooth(aes(x=age, y=creatclear))
```

As we can see, the difference between the line of best fit and the 40 year old is much greater than the difference between the line of best fit and the 60 year old. Therefore, the 40 year old creatclear is healthier for their age.

# Problem 2

### Part A
```{r}
covid = read.csv("covid.csv")
# Italy
ItalyC = subset(covid, country == "Italy")
ItalyB = do(10000)*lm(log(deaths) ~ days_since_first_death, data=resample(ItalyC))
confint(ItalyB, level = 0.95)
```
Growth rate: 0.183. 95% CI of [0.160, 0.208]. Doubling time: $T_d = ln(2) /\ 0.183\approx 3.8 \text{ days}$  

### Part B
```{r}
# Spain
SpainC = subset(covid, country == "Spain")
SpainB = do(10000)*lm(log(deaths) ~ days_since_first_death, data=resample(SpainC))
confint(SpainB, level = 0.95)
```
Growth rate: 0.276. 95% CI of [0.235, 0.317]. Doubling time: $T_d = ln(2) /\ 0.276 \approx 2.5 \, \text{days}$  

### Part C
```{r}
# line graph
ggplot(covid) + 
  geom_line(aes(x = days_since_first_death, y = deaths, col = country))
```

# Problem 3

```{r}
milk = read.csv("milk.csv")
milkB = do(10000)*lm(log(sales) ~ log(price), data=resample(milk))
confint(milkB, level = 0.95)
```
Estimated price elasticity is -1.619 with a 95% CI of [-1.772, -1.459].

To estimate the price elasticity, I used a log-log regression model (log(sales) ~ log(price)) that directly estimates the elasticity coefficient in the power-law model. I drew 10,000 bootstrap resamples of the milk data to account for statistical uncertainty, using the do() and resample() functions of the mosaic package. The coefficient of log(price) is the price elasticity of demand. The elasticity estimate of -1.619 implies that milk demand is elastic - a 1% price increase leads to roughly a 1.62% decrease in quantity demanded. Since the confidence interval is all less than -1, we are 95% confident that milk demand is elastic, i.e., consumers are reasonably sensitive to price changes.
