---
title: "Homework 2 - SDS315"
author: "Adam Jose - avj423"
date: "01/27/2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

# Problem 1: Beauty or Not, in the Classroom
## Part A: Histogram of Course Evalution Scores

```{r}
library(ggplot2)
library(dplyr)

# Load data
profes <- read.csv("profs.csv")

# Create histogram
ggplot(profes, aes(x = eval)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Course Evaluation Scores",
       x = "Evaluation Score",
       y = "Frequency") +
  theme_minimal()
```
The histogram shows the distribution of course evaluation scores. Most scores are concentrated between 3.5 and 4.5, indicating that the majority of instructors receive above-average evaluations.

## Part B: Boxplot of Evaluation Scores by Native English Speaker
```{r}
# Create boxplot
ggplot(profes, aes(x = native, y = eval, fill = native)) +
  geom_boxplot() +
  labs(title = "Course Evaluation Scores by Native English Speaker Status",
       x = "Native English Speaker",
       y = "Evaluation Score") +
  theme_minimal()
```
The boxplot compares the distribution of course evaluation scores for native and non-native English speakers. There is no significant difference in evaluation scores between the two groups.

## Part C: Faceted Histogram by Gender
```{r}
# Create faceted histogram
ggplot(profes, aes(x = eval)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  facet_wrap(~ gender, nrow = 2) +
  labs(title = "Distribution of Course Evaluation Scores by Gender",
       x = "Evaluation Score",
       y = "Frequency") +
  theme_minimal()
```
The faceted histogram compares the distribution of course evaluation scores for male and female instructors. Both genders show similar distributions, with most scores between 3.5 and 4.5.

## Part D: Scatterplot of Attractiveness vs. Evaluation Scores

```{r}
# Create scatterplot
ggplot(profes, aes(x = beauty, y = eval)) +
  geom_point(alpha = 0.5, color = "blue") +
  labs(title = "Association Between Professor Attractiveness and Course Evaluations",
       x = "Attractiveness Score",
       y = "Evaluation Score") +
  theme_minimal()
```
The scatterplot shows a weak positive association between a professor's physical attractiveness and their course evaluation scores. However, the relationship is not strong.

# Problem 2: Bike Sharing
## Plot A: Average Hourly Bike Rentals

```{r}
# Load data
bikeshare <- read.csv("bikeshare.csv")

# Calculate average hourly rentals
avg_hourly <- bikeshare %>%
  group_by(hr) %>%
  summarise(avg_rentals = mean(total))

# Create line graph
ggplot(avg_hourly, aes(x = hr, y = avg_rentals)) +
  geom_line(color = "blue") +
  labs(title = "Average Hourly Bike Rentals",
       x = "Hour of the Day",
       y = "Average Rentals") +
  theme_minimal()
```
The line graph shows the average number of bike rentals per hour. Rentals peak during morning and evening rush hours, with the highest demand around 8 AM and 5 PM.

## Plot B: Faceted Line Graph by Working Day
```{r}
# Calculate average hourly rentals by working day
avg_hourly_workingday <- bikeshare %>%
  group_by(hr, workingday) %>%
  summarise(avg_rentals = mean(total))

# Create faceted line graph
ggplot(avg_hourly_workingday, aes(x = hr, y = avg_rentals, color = factor(workingday))) +
  geom_line() +
  facet_wrap(~ workingday, nrow = 2) +
  labs(title = "Average Hourly Bike Rentals by Working Day",
       x = "Hour of the Day",
       y = "Average Rentals",
       color = "Working Day") +
  theme_minimal()
```
The faceted line graph compares average hourly bike rentals on working days versus non-working days. On working days, rentals peak during commute hours, while on non-working days, rentals are more evenly distributed throughout the day.

## Plot C: Faceted Bar Plot for 9 AM Rentals by Weather

```{r}
# Filter data for 9 AM
bikeshare_9am <- bikeshare %>%
  filter(hr == 9)

# Calculate average rentals by weather and working day
avg_9am_weather <- bikeshare_9am %>%
  group_by(weathersit, workingday) %>%
  summarise(avg_rentals = mean(total))

# Create faceted bar plot
ggplot(avg_9am_weather, aes(x = factor(weathersit), y = avg_rentals, fill = factor(workingday))) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ workingday) +
  labs(title = "Average 9 AM Bike Rentals by Weather and Working Day",
       x = "Weather Situation",
       y = "Average Rentals",
       fill = "Working Day") +
  theme_minimal()
```
The bar plot shows average bike rentals during the 9 AM hour, faceted by working day and weather situation. Rentals are higher on working days, and poor weather (e.g., rain or snow) reduces ridership.

# Problem 3: Capital Metro UT Ridership
## Plot 1: Average Boardings by Hour, Day, and Month

```{r}
# Load data
capmetro_UT <- read.csv("capmetro_UT.csv")

# Reorder categorical variables
capmetro_UT <- capmetro_UT %>%
  mutate(day_of_week = factor(day_of_week, levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")),
         month = factor(month, levels = c("Sep", "Oct", "Nov")))

# Calculate average boardings
avg_boardings <- capmetro_UT %>%
  group_by(hour_of_day, day_of_week, month) %>%
  summarise(avg_boardings = mean(boarding))

# Create faceted line graph
ggplot(avg_boardings, aes(x = hour_of_day, y = avg_boardings, color = month)) +
  geom_line() +
  facet_wrap(~ day_of_week) +
  labs(title = "Average Boardings by Hour, Day, and Month",
       x = "Hour of the Day",
       y = "Average Boardings",
       color = "Month") +
  theme_minimal()
```
The faceted line graph shows average boardings by hour of the day, day of the week, and month. Peak boardings occur around 8 AM and 5 PM on weekdays, with lower boardings on weekends. Boardings in September are lower on Mondays, possibly due to the start of the semester, while boardings in November are lower on Wednesdays, Thursdays, and Fridays, possibly due to holidays.

## Plot 2: Boardings vs. Temperature by Hour and Weekend
```{r}
# Create faceted scatter plot
ggplot(capmetro_UT, aes(x = temperature, y = boarding, color = weekend)) +
  geom_point(alpha = 0.5) +
  facet_wrap(~ hour_of_day) +
  labs(title = "Boardings vs. Temperature by Hour and Weekend",
       x = "Temperature (F)",
       y = "Boardings",
       color = "Weekend") +
  theme_minimal()
```
The scatter plot shows boardings versus temperature, faceted by hour of the day and colored by weekend status. Temperature has a minimal effect on boardings when controlling for hour and weekend status.

# Problem 4: Wrangling the Billboard Top 100
## Part A: Top 10 Most Popular Songs

```{r}
# Load data
billboard <- read.csv("billboard.csv")

# Calculate total weeks on chart
top_songs <- billboard %>%
  group_by(performer, song) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(10)

# Print table
knitr::kable(top_songs, caption = "Top 10 Most Popular Songs by Total Weeks on Billboard Top 100")
```
The table lists the top 10 most popular songs since 1958, measured by the total number of weeks spent on the Billboard Top 100 chart.

## Part B: Musical Diversity Over Time
```{r}
# Filter out 1958 and 2021
billboard_filtered <- billboard %>%
  filter(year != 1958 & year != 2021)

# Calculate unique songs per year
unique_songs <- billboard_filtered %>%
  group_by(year) %>%
  summarise(unique_songs = n_distinct(song))

# Create line graph
ggplot(unique_songs, aes(x = year, y = unique_songs)) +
  geom_line(color = "blue") +
  labs(title = "Musical Diversity Over Time",
       x = "Year",
       y = "Number of Unique Songs") +
  theme_minimal()
```
The line graph shows the number of unique songs appearing on the Billboard Top 100 each year. Musical diversity has increased over time, with more unique songs appearing in recent years.

## Part C: Ten-Week Hits by Artist
```{r}
# Calculate ten-week hits
ten_week_hits <- billboard %>%
  group_by(performer, song) %>%
  summarise(weeks_on_chart = n()) %>%
  filter(weeks_on_chart >= 10) %>%
  group_by(performer) %>%
  summarise(count = n()) %>%
  filter(count >= 30)

# Create bar plot
ggplot(ten_week_hits, aes(x = reorder(performer, count), y = count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  coord_flip() +
  labs(title = "Ten-Week Hits by Artist",
       x = "Artist",
       y = "Number of Ten-Week Hits") +
  theme_minimal()
```
The bar plot shows the number of "ten-week hits" (songs that appeared on the Billboard Top 100 for at least 10 weeks) for the 19 artists with at least 30 such hits. The artists are ordered by the number of hits.














